\documentclass[11pt,a4paper,twoside,openright]{book}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{fullpage}
\usepackage[usenames,dvipsnames]{color}
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{wrapfig}							% allows wrapping text around figures
\usepackage{multicol}
%\usepackage{cite}
%\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\titleformat{\section}[block]{\large\bfseries}{\thesection}{1em}{}
\usepackage{framed}
\usepackage{setspace}
\def\beq{\begin{equation*}}
\def\eeq{\end{equation*}}
\def\beqn{\begin{matrix}}
\def\eeqn{\end{matrix}}

\setlength{\marginparwidth}{0pt}

 % Define commands to assure consistent treatment throughout document
% \newcommand{\eqnref}[1]{(\ref{#1})}
 \newcommand{\class}[1]{\texttt{#1}}
 \newcommand{\package}[1]{\texttt{#1}}
 \newcommand{\file}[1]{\texttt{#1}}
 \newcommand{\BibTeX}{\textsc{Bib}\TeX}
\renewcommand{\thesection}{\Roman{section}.}
\renewcommand{\thesubsection}{\arabic{subsection}.}
\renewcommand{\thesubsubsection}{\alph{subsubsection}.}

\begin{document}

\begin{center}
\begin{LARGE}{\bf Aircraft Flight Mechanics}\end{LARGE}\\
\large
\vspace{22 mm}
\begin{singlespace}
   A Brief Textbook Presented to the \\ 
   Student Body of the University of South Alabama \\
\vspace{22 mm}
   by\\
\vspace{22 mm}
  Carlos Montalvo \\
\vspace{22 mm}
\end{singlespace}
{\itshape University of South Alabama}
{\bf Copyright $\copyright$ Carlos Jos\'{e} Montalvo 2015}
\end{center}

\newpage

\section{Particle Dynamics}
For this formulation we start with Newton's Second Law with no
approximations.
\begin{equation}
\sum\limits_{i=0}^N \vec{F}_{ji} = \frac{d\vec{p}_j}{dt}
\end{equation}
where $\vec{p}_j$ is the momentum of a particle. $\vec{F}_{ji}$ is a
force on the particle. The statement above states that sum of all
forces on a particle is equal to the time rate of change of
momentum. 

\subsection{Linear Dynamics for Systems of Particles}
If two particles are then considered the equation can
be written for both particles.
\begin{equation}
\begin{matrix}
\sum\limits_{i=0}^N \vec{F}_{1i} + \vec{f}_{12} = \frac{d\vec{p}_1}{dt} &
\sum\limits_{i=0}^N \vec{F}_{2i} + \vec{f}_{21} = \frac{d\vec{p}_2}{dt} 
\end{matrix}
\end{equation}
Note that the forces $\vec{f}_{12}$ and $\vec{f}_{21}$ are internal forces
experienced by each particle exerted on each other since they are
rigidly connected. Newton's Third Law states that for every action
there is an equal and opposite reaction. That is, $\vec{f}_{12} = -\vec{f}_{21}$. Thus, if both equations are added the following equation is
created
\begin{equation}
\sum\limits_{j=0}^P \sum\limits_{i=0}^N \vec{F}_{ji} =
\sum\limits_{j=0}^P \frac{d\vec{p}_j}{dt}
\end{equation}
where P is the number of particles. Typically the double summation
in F is written just as $\vec{F}$. 

\subsection{Rotational Dynamics for Systems of Particles}

Note that by construction, a system of particles rigidly connected can
now rotate about a center point. The center of mass of a system of
particles can be defined using the relationship below
\begin{equation}
\vec{r}_C = \frac{1}{m}\sum\limits_{j=0}^P m_j\vec{r}_{j}
\end{equation}
where
\begin{equation}
m = \sum\limits_{j=0}^P m_j
\end{equation}
This vector can then be used to create rotational dynamics starting
with the linear dynamics.
\begin{equation}
\sum\limits_{j=0}^P \sum\limits_{i=0}^N {\bf S}(\vec{r}_{Cj}) \vec{F}_{ji}  = \vec{M}_C = \sum\limits_{j=0}^P {\bf
  S}(\vec{r}_{Cj}) \frac{d\vec{p}_j}{dt}
\end{equation}
where ${\bf S}(\vec{r}_{Cj})$ is the skew symmetric matrix of the
vector from the center of mass to the jth particle which results in a cross product.
\section{Rigid Bodies}
At this point, many assumptions are made about the
system of particles.
\begin{enumerate}
\item The mass of each particle or rigid body is constant.
\item The Earth is assumed to be flat and not rotating. This allows
  the Earth to be used as an inertial frame.
\item The rigid body is not flexible and does not change shape. That
  is, the time rate of change of the magnitude of a vector
  $\vec{r}_{PQ}$ is zero for any arbitrary points P and Q attached to
  the rigid body.
\end{enumerate}
\subsection{Linear Dynamics}
Using all of these simplifications, the momentum term on the right can
be simplified to 
\begin{equation}
\sum\limits_{j=0}^P \vec{p}_j = m \vec{v}_{C/I}
\end{equation}
The derivation of the term above starts by deriving the position of
the center of mass as the following equation.
\begin{equation}
\vec{r}_{j} = \vec{r}_C + \vec{r}_{Cj}
\end{equation}
Taking one derivative results in the following equation
\begin{equation}
\vec{v}_{j/I} = \vec{v}_{C/I} + \frac{{}^Bd \vec{r}_{Cj}}{dt} +
{\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}
\end{equation}
where ${\bf S}(\vec{\omega}_{B/I})$ is the skew symmetric matrix of the
angular velocity vector which results in a cross product. This
equation comes from the derivative transport theorem. Since the body
is a rigid body the term $\frac{{}^Bd \vec{r}_{Cj}}{dt}=0$ resulting
in the equation below
\begin{equation}
\vec{v}_{j/I} = \vec{v}_{C/I} + {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}
\end{equation}
which any dynamicist knows as the equation for two points fixed on a
rigid body. This equation can then be substituted into the equation
for momentum such that.
\begin{equation}
\sum\limits_{j=0}^P \vec{p}_j =  \sum\limits_{j=0}^P m_j \left(\vec{v}_{C/I}
+ {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}\right)
\end{equation}
The first term reduces to 
\begin{equation}
\sum\limits_{j=0}^P m_j \vec{v}_{C/I} =  \vec{v}_{C/I}
\sum\limits_{j=0}^P m_j =  m \vec{v}_{C/I} 
\end{equation}
the second term reduces to zero since the sum of all particles from
the center of mass is by definition the center of mass and thus zero.
\begin{equation}
\sum\limits_{j=0}^P {\bf S}(\vec{\omega}_{B/I}) m_j\vec{r}_{Cj} =
{\bf S}(\vec{\omega}_{B/I})\sum\limits_{j=0}^P m_j\vec{r}_{Cj} = 0
\end{equation}
Plugging this result for momentum into Newton's equation of motion
yields. This is typically called Newton-Euler equations of motion.
\begin{equation}
\vec{F}_C = m \left(\frac{{}^Bd \vec{v}_{C/I}}{dt} +
{\bf S}(\vec{\omega})_{B/I} \vec{v}_{C/I} \right)
\end{equation}
\subsection{Rotational Dynamics}
Plugging in the expression for two points fixed on a rigid body
results in a much different expression. First let's expand the
rotational dynamic equations of particles using the assumptions made
for a rigid body.
\begin{equation}
\vec{M}_C = \frac{d}{dt}\sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj}) m_j\vec{v}_{j/I}
\end{equation}
Then the equation of two points fixed on a rigid body can be
introduced to obtain the following equation
\begin{equation}
\vec{M}_C = \frac{d}{dt}\sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj}) m_j\left(\vec{v}_{C/I} + {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj}\right)
\end{equation}
expanding this into two terms yields
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}) {\bf S}(\vec{\omega}_{B/I}) \vec{r}_{Cj} + \sum\limits_{j=0}^P {\bf S}(\vec{r}_{Cj})m_j\vec{v}_{C/I}\right)
\end{equation}
To simplify this further a useful equality is used for cross
products. That is ${\bf S}(\vec{a})\vec{b}=-{\bf
  S}(\vec{b})\vec{a}$. The equation above then changes to
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\left(-\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})\right)\vec{\omega}_{B/I} - {\bf S}(\vec{v}_{C/I})
\sum\limits_{j=0}^P \vec{r}_{Cj}m_j\right)
\end{equation}
Notice, that parentheses were placed around the first term to isolate
the angular velocity. This is because the angular velocity is constant
across the system of particles. The term on the right has also been
altered slightly to isolate the fact that the velocity of the center
of mass is independent of the system of particles. With the equation
in this form it is easy to see that the term on the right is zero
because it is the definition of the center of mass. The equation then
reduces to 
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left(\sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})^T\right)\vec{\omega}_{B/I} 
\end{equation}
Notice again that minus sign has been removed. The skew symmetric
matrix has an interesting property where the transpose is equal to the
negative of the original matrix. The term in brackets is a well known
value for rigid bodies and is known as the moment of inertia for rigid
bodies. 
\begin{equation}
{\bf I}_C =  \sum\limits_{j=0}^P m_j{\bf S}(\vec{r}_{Cj}){\bf
  S}(\vec{r}_{Cj})^T
\end{equation}
This results in the kinematic equations of motion for rigid bodies to
the simple equation below.
\begin{equation}
\vec{M}_C =  \frac{d}{dt}\left({\bf I}_C\vec{\omega}_{B/I} \right)
\end{equation}
With the equation in this form it is finally possible to carry out the
derivative
\begin{equation}
\vec{M}_C = \frac{{}^Bd ({\bf I}_C\vec{\omega}_{B/I})}{dt} + {\bf
  S}(\vec{\omega}_{B/I}){\bf I}_C\vec{\omega}_{B/I}
\end{equation}
The first term requires the chain rule to perform the derivative
however the body frame derivative of the moment of inertia matrix is
simply zero. Therefore the equation can simply be written as
\begin{equation}
\vec{M}_C = {\bf I}_C\frac{{}^Bd (\vec{\omega}_{B/I})}{dt} + {\bf
  S}(\vec{\omega}_{B/I}){\bf I}_C\vec{\omega}_{B/I}
\end{equation}
\section{Aircraft Convention}
Aircraft convention involves using the Newton-Euler equations of
motion to describe the aircraft. Typically the position of the
aircraft is written as 
\begin{equation}
{\bf C}_I(\vec{r}_C) = \begin{Bmatrix} x \\ y \\ z \end{Bmatrix}
\end{equation}
The derivative of the position vector is the velocity vector is then
written as
\begin{equation}
{\bf C}_I(\vec{v}_{C/I}) = \begin{Bmatrix} \dot{x} \\ \dot{y} \\ \dot{z} \end{Bmatrix}
\end{equation}
However, body frame coordinates are typically used to
describe the velocity vector such that
\begin{equation}
{\bf C}_B(\vec{v}_{C/I}) = \begin{Bmatrix} u \\ v \\ w \end{Bmatrix}
\end{equation}
In order to relate the body frame components of the velocity vector
the inertial frame coordinates a transformation matrix is used. The
transformation from the inertial frame to the body frame involves
three unique rotations. The first is a rotation about the z-axis such
that
\begin{equation}
{\bf C}_A(\vec{v}_{C/I}) = \begin{bmatrix} cos(\psi) & sin(\psi) & 0
  \\ -sin(\psi) & cos(\psi) & 0 \\ 0 & 0 & 1 \end{bmatrix} {\bf C}_I(\vec{v}_{C/I})
\end{equation}
this rotation is called the yaw or heading rotation. From here the
intermediate frame is rotated about the y-axis such that
\begin{equation}
{\bf C}_{NR}(\vec{v}_{C/I}) = \begin{bmatrix} cos(\theta) & 0 &
  -sin(\theta) \\ 0 & 1 & 0 \\ sin(\theta) & 0 & cos(\theta) \end{bmatrix} {\bf C}_A(\vec{v}_{C/I})
\end{equation}
this rotation is called the pitch angle rotation. Finally the no roll
frame is rotated through the x-axis such that
\begin{equation}
{\bf C}_{B}(\vec{v}_{C/I}) = \begin{bmatrix} 1 & 0 & 0 \\ 0 & cos(\phi) & sin(\phi)
  \\ 0 & -sin(\phi) & cos(\phi) \end{bmatrix} {\bf C}_{NR}(\vec{v}_{C/I})
\end{equation}
Putting all the equations together yields
\begin{equation}\label{e:xyzdot}
\begin{Bmatrix} \dot{x} \\ \dot{y} \\ \dot{z}   \end{Bmatrix} = [\textbf{T}_{IB}]
\begin{Bmatrix} u \\ v \\ w \end{Bmatrix}
\end{equation}
where 
\begin{equation}\label{e:TIB}
\textbf{T}_{IB} = \begin{bmatrix} c_{\theta}c_{\psi} &
s_{\phi}s_{\theta}c_{\psi}-c_{\phi}s_{\psi} &
c_{\phi}s_{\theta}c_{\psi} + s_{\phi}s_{\psi} \\ c_{\theta}s_{\psi} &
s_{\phi}s_{\theta}s_{\psi} + c_{\phi}c_{\psi} &
c_{\phi}s_{\theta}s_{\psi} - s_{\phi}c_{\psi} \\
-s_{\theta} & s_{\phi}c_{\theta} & c_{\phi}c_{\theta}
\end{bmatrix}
\end{equation}
Standard shorthand notation is used for trigonometric
functions: $ cos(\alpha) \equiv c_{\alpha} $ , $ sin(\alpha) \equiv
s_{\alpha} $ , and $ tan(\alpha) \equiv t_{\alpha} $. These three
angles are known as the standard Euler angle rotation sequence of an
aircraft in free flight. The angular
velocity of a body is typically written as 
\begin{equation}
{\bf C}_B(\vec{\omega}_{B/I}) = \begin{Bmatrix} p \\ q
  \\ r \end{Bmatrix} = p \hat{I}_B + q \hat{J}_B + r \hat{K}_B
\end{equation}
There are no inertial components for the angular velocity
vector. However, a relationship can be derived relating the
derivatives of the Euler angles. The angular velocity can be written
in vector form such that
\begin{equation}
\vec{\omega}_{B/I} = \dot{\psi} \hat{K}_A + \dot{\theta} \hat{J}_{NR} +
\dot{\phi} \hat{I}_B
\end{equation}
relating the unit vectors $\hat{K}_A$ and $\hat{J}_{NR}$ to the body
frame using the planar rotation matrices results in the equation
below. Note that NR is denoted as the ``No-Roll" frame.
\begin{equation}\label{e:ptpdot}
\begin{Bmatrix} \dot{\phi} \\ \dot{\theta} \\ \dot{\psi} \end{Bmatrix}
= [\textbf{H}]
\begin{Bmatrix} p\\q\\r\end{Bmatrix}
\end{equation}
where
\begin{equation}
\textbf{H}=\begin{bmatrix} 1 & s_{\phi}t_{\theta} & c_{\phi}t_{\theta} \\ 0 &
c_{\phi} & -s_{\phi} \\ 0 & s_{\phi}/c_{\theta} &
c_{\phi}/c_{\theta} \end{bmatrix}
\end{equation}
with all this information the Newton-Euler equations of motion can be
used to form the equation below.
\begin{equation}\label{e:uvwdot} 
\begin{Bmatrix} \dot{u} \\ \dot{v} \\ \dot{w} \end{Bmatrix} = 
\frac{1}m \begin{Bmatrix} X\\Y\\Z\end{Bmatrix}-\begin{bmatrix} 0 & -r
& q \\ r & 0 & -p \\ -q & p & 0 \end{bmatrix} \begin{Bmatrix} u\\v\\w\end{Bmatrix}
\end{equation}
Using the rotational dynamic equations for rigid bodies, the equation
for the derivative of angular velocity can be found as
\begin{equation}\label{e:pqrdot} 
\begin{Bmatrix} \dot{p} \\ \dot{q} \\ \dot{r} \end{Bmatrix} = {\bf I}_C^{-1}\left(
\begin{Bmatrix} L\\M\\N\end{Bmatrix}-\begin{bmatrix} 0 & -r
& q \\ r & 0 & -p \\ -q & p & 0 \end{bmatrix} {\bf I}_C\begin{Bmatrix} p\\q\\r\end{Bmatrix}\right)
\end{equation}
Note that standard aircraft forces and moments are applied to the
body. The forces are typically written as X,Y and Z while the moments
are given as L,M and N. They can be written in component form using
the equations below.
\begin{equation}
{\bf C}_B(\vec{F}_{C}) = \begin{Bmatrix} X \\ Y
  \\ Z \end{Bmatrix} = X \hat{I}_B + Y \hat{J}_B + Z \hat{K}_B
\end{equation}
\begin{equation}
{\bf C}_B(\vec{M}_{C}) = \begin{Bmatrix} L \\ M
  \\ N \end{Bmatrix} = L \hat{I}_B + M \hat{J}_B + N \hat{K}_B
\end{equation}

\section{Forces on Aircraft}
To form the sum of the forces on an aircraft the assumptions are made
that only gravity and aerodynamic forces act. For rockets, a
propulsion model can be added. Still the gravitational force is shown
below. 
\subsection{Gravity}
The weight contribution is assumed to be a constant force applied to
the aircraft. The equation below is the gravitational force applied in
the body frame.
\begin{equation}\label{e:wforce}
\begin{Bmatrix} X_W \\ Y_W \\ Z_W \end{Bmatrix} = mg \begin{Bmatrix}
-s_{\theta} \\ s_{\phi}c_{\theta} \\ c_{\phi}c_{\theta} \end{Bmatrix}
\end{equation}
\subsection{Aerodynamics}
Aircraft aerodynamics are written using a taylor series expansion
about a trim point. That is, the aerodynamic forces are given by
\begin{equation}
\vec{F} = \vec{F}_0 + \frac{\partial \vec{F}}{\partial \vec{x}}(\vec{x}-\vec{x}_0)
\end{equation}
where $\vec{x} = [x,y,z,\phi,\theta,\psi,u,v,w,p,q,r]^T$. The partial
derivative is thus expanded such that
\begin{equation}
\frac{\partial \vec{F}}{\partial \vec{x}} = \begin{bmatrix} \frac{\partial
    \vec{F}}{\partial x} & \frac{\partial \vec{F}}{\partial y} & ... &
  \frac{\partial \vec{F}}{\partial r} \end{bmatrix}
\end{equation}
To find all of the partial derivative the forces are first written
using a combination of dynamic pressure and coefficients that are
functions of geometry and Reynolds number rather than speed, pressure
and size. A general lift force can be written using the equation below
\begin{equation}
L = \frac{1}2\rho {V_{\infty}}^2 S C_L
\end{equation}
where $\rho$ is the atmospheric density, $V_{\infty}$ is the
free-stream velocity, S is the planform area of the wing and $C_L$ is
the lift coefficient. 
\begin{equation}\label{e:vtotal}
  V_{\infty} = \sqrt{{u_a}^2 + {v_a}^2 + {w_a}^2}
\end{equation}
The subscript 'a' above denotes the velocity of the aircraft plus the
atmospheric disturbance. 
\begin{equation}\label{e:atm}
\begin{Bmatrix} u_a \\ v_a \\ w_a \end{Bmatrix} =
\begin{Bmatrix} u \\ v \\ w \end{Bmatrix} +
{\textbf{T}^T}_{IB} \begin{Bmatrix}
 V_x \\ V_y \\ V_z \end{Bmatrix}
\end{equation}
Note that the dynamic pressure is different for
a rocket or projectile. A similar expression can be created for a
generic moment such that
\begin{equation}
M = \frac{1}2\rho {V_{\infty}}^2 S \bar{c} C_M
\end{equation}
where $\bar{c}$ is the mean chord of the aircraft. The dynamic pressure $q_{\infty} =
\frac{1}2\rho {V_{\infty}}^2 S$ can be used to non-dimensionalize the forces, thus $L/q_{\infty} = C_L$. This means
that the equation involving partial derivatives can be written as
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix} \frac{\partial
    \vec{C}_F}{\partial x} & \frac{\partial \vec{C}_F}{\partial y} & ... &
  \frac{\partial \vec{C}_F}{\partial r} \end{bmatrix}
\end{equation}
If the vector is then expanded to include the components of the vector
$\vec{F}$ the partial derivatives expand to
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix}
  \frac{\partial C_X}{\partial x} & \frac{\partial C_X}{\partial y} &
  ... & \frac{\partial C_x}{\partial r} \\ \frac{\partial C_Y}{\partial x} & \frac{\partial C_Y}{\partial y} &
  ... & \frac{\partial C_Y}{\partial r} \\ \frac{\partial C_Z}{\partial x} & \frac{\partial C_Z}{\partial y} &
  ... & \frac{\partial C_Z}{\partial r} \end{bmatrix}
\end{equation}
shorthand can be adopted for the forces above such that
$\frac{\partial C_Y}{\partial x} = C_{Yx}$. Using this shorthand the
equation above can be written as.
\begin{equation}
\frac{\partial \vec{C}_F}{\partial \vec{x}} = \begin{bmatrix}
  C_{Xx} & C_{Xy} &
  ... & C_{Xr} \\ C_{Yx} & C_{Yy} &
  ... & C_{Yr} \\ C_{Zx} & C_{Zy} &
  ... & C_{Zr} \end{bmatrix}
\end{equation}
The coefficients listed above are standard coefficients that all
aircraft have. A similar matrix can be formulated for the moments on
an aircraft. When system identifying an aircraft all of these
coefficients may be determined. However, many of these terms are
zero. For example, all coefficients with respect to x y and z are
zero. That is, $C_{Xx} = C_{Yx} = ... C_{Nx} = C_{Xy} = ... C_{Nz} =
0$. Other coefficients can be set to zero as well. 
\subsubsection{Aircraft Aerodynamics}
For aircraft, some further simplifications are made. Some of the
coefficients defined above are combined to be written as functions of the
angle of attack($\alpha$) and sideslip($\beta$).
\begin{equation}\label{e:aoa}
\alpha = tan^{-1}\left(\frac{w_a} {u_a} \right)
\end{equation}
\begin{equation}\label{e:beta}
\beta = sin^{-1}\left(\frac{v_a}{V_{\infty}}\right)
\end{equation}
Transforming the equations into these formulations gives rise to
coefficients such as $C_{L\alpha}$ which is the change in lift as a
function of angle of attack and $C_{Y\beta}$ which is the change in
Y-Force as a function of sideslip. Using all of the coefficients
defined above taking into account the change to lift and drag, the body
aerodynamic force is calculated using the equation below.
\begin{equation}\label{e:aforce}
\begin{Bmatrix} X_A \\ Y_A \\ Z_A \end{Bmatrix} = \frac{1}2\rho
{V_{\infty}}^2 S \begin{Bmatrix}
C_Ls_{\alpha}-C_Dc_{\alpha}+C_{x_{\delta_t}} \delta_t
\\ C_{y{\beta}}{\beta}+C_{y{\delta}_r}{\delta}_r+C_{yp}\frac{pb}{2V_{\infty}}
+ C_{yr}\frac{rb}{2V_{\infty}} \\ -C_Lc_{\alpha}-C_Ds_{\alpha} \end{Bmatrix}
\end{equation}
Where the lift and drag coefficients are:
\begin{equation}\label{e:liftdrag}
\begin{Bmatrix} C_L \\ C_D \end{Bmatrix} = \begin{Bmatrix} C_{L0} +
C_{L\alpha}\alpha + C_{Lq}\frac{q\bar{c}}{2V_{\infty}} + C_{L\delta_e}\delta_e \\   C_{D0} +
C_{D\alpha}\alpha^2 \end{Bmatrix}
\end{equation}
The body aerodynamic moment is also computed using an aerodynamic expansion.
\begin{equation}\label{e:LMN}
\begin{Bmatrix} L_A \\ M_A \\ N_A \end{Bmatrix} = \frac{1}2\rho
{V_{\infty}}^2 S \bar{c} \begin{Bmatrix} C_{l\beta}\beta + C_{lp}\frac{pb}{2V_{\infty}} + C_{lr}\frac{rb}{2V_{\infty}} + C_{l\delta_a}{\delta_a} + C_{l\delta_r}{\delta_r}
\\  C_{m0} + C_{m\alpha}\alpha + C_{mq}\frac{q\bar{c}}{2V_{\infty}}+ C_{m\delta_e}\delta_e  \\ C_{np}\frac{pb}{2V_{\infty}} + C_{n\beta}\beta + C_{nr}\frac{rb}{2V_{\infty}} + C_{n\delta_a}\delta_a + C_{n\delta_r}\delta_r \end{Bmatrix}
\end{equation}
The aerodynamic coefficients in equations
(\ref{e:aforce}), (\ref{e:liftdrag}) and (\ref{e:LMN}) can be obtained from
flight data, aerodynamic modeling and windtunnel tests. Notice that
the only coefficients remaining are coefficients from angle of attack,
sideslip and angular velocities. Furthermore, the coefficients for
angular velocities are also non-dimensionalized by terms such as
$b/(2V_{\infty})$ where $b$ is the wingspan of the aircraft and
$\bar{c}$ is the mean chord of the aircraft. These terms are
introduced to fully non-dimensionalize the coefficients. Notice, as
well that four extra terms were also introduced. These will 
be discussed in more detail in the control section however the four
terms are the aileron control surface $\delta_a$, the elevator
control surface $\delta_e$, the rudder control surface $\delta_r$ and
the thrust control value $\delta_t$. 

\subsubsection{Projectile Aerodynamics}

To fully define the projectile aerodynamics some more assumptions are
made about the projectile.
\begin{enumerate}
\item The projectile is axially symmetric
\item The aerodynamic forces are not necessarily formulated at the center
  of mass
\item The projectile has the potential to be spinning rapidly thus
  interacting with the surrounding atmosphere
\end{enumerate}
For a projectile the dynamic pressure is written as
\begin{equation}
Q = \frac{\pi}{8}\rho V_{\infty}^2d^2 
\end{equation}
The aerodynamic forces on the projectile are modeled using taylor series
ballistic expansions with known coefficients similar to the aircraft
model only slightly different assumptions are made given the dynamics
of the projectile. The subscripts in the equation below stand for
steady and unsteady aerodynamics.
\begin{equation}\label{e:aeroF}
\begin{Bmatrix} X_A \\ Y_A\\ Z_A \end{Bmatrix} = \begin{Bmatrix}
  X_{SA} \\ Y_{SA}\\ Z_{SA} \end{Bmatrix} + \begin{Bmatrix}
  X_{UA} \\ Y_{UA}\\ Z_{UA} \end{Bmatrix} = 
Q \begin{Bmatrix} -C_{X_0} - C_{X_2}\frac{v^2+w^2}{V^2}\\
-C_{Y_\beta}\frac{v}{V}\\
-C_{N_\alpha}\frac{w}{V}
\end{Bmatrix} + Q \begin{Bmatrix} 0
  \\ C_{Y_{p\alpha}}\frac{w}{V}\frac{pd}{2V} \\ C_{Z_{p\alpha}}\frac{v}{V}\frac{pd}{2V}\end{Bmatrix}
\end{equation}
In this equation, $Q$ is the dynamic pressure, $d$ is the aerodynamic
reference area, $C_{X_0}$ is the zero-yaw axial force
coefficient,$C_{X_2}$ is the yaw-squared axial force coefficient,
$C_{N_\alpha}$ is the normal force derivative coefficient,
$C_{Y_{p\alpha}}$ is the Magnus force coefficient, and
$V=\sqrt{u^2+v^2+w^2}$ is the total velocity of the projectile. 
The aerodynamic moments acting on the projectile are the pitching,
pitch damping, Magnus, and roll damping moments. Pitching and Magnus
moments are given by taking the cross product of the normal and Magnus
forces given in (\ref{e:aeroF})  with the position vector from the
center of mass to the center of pressure and location of Magnus force,
respectively. The total aerodynamic moments are given in
Eqn. (\ref{e:aeroM}). 
\begin{equation}\label{e:aeroM}
\begin{Bmatrix} L_A \\ M_A \\ N_A \end{Bmatrix}=
{\bf S}_B(\vec{r}_{CG,COP})\begin{Bmatrix}
  X_{SA} \\ Y_{SA}\\ Z_{SA} \end{Bmatrix} + {\bf S}_B(\vec{r}_{CG,MCOP})\begin{Bmatrix}
  X_{UA} \\ Y_{UA}\\ Z_{UA} \end{Bmatrix} + 
Qd \begin{Bmatrix}
C_{l_p} \frac{pd}{2V} \\
C_{m_q} \frac{qd}{2V}\\
C_{n_r} \frac{rd}{2V} \end{Bmatrix}
\end{equation}
Here, ${\bf S}_B(\vec{r}_{CG,COP})$ is the skew-symmetric operator
acting on the position vector from the center of mass to the center of
pressure expressed in the projectile body frame. Furthermore, ${\bf S}_B(\vec{r}_{CG,MCOP})$ is the skew-symmetric operator
acting on the position vector from the center of mass to the magnus
center of pressure expressed in the projectile body frame. Typically
the center of mass is defined from the rear of the projectile such
that 
\begin{equation}
{\bf C}_B(\vec{r}_{CG}) = \begin{Bmatrix} SL_{CG} \\ BL_{CG} \\ WL_{CG} \end{Bmatrix}
\end{equation}
Similarly, the center of pressure is defined from the rear of the
projectile such that
\begin{equation}
{\bf C}_B(\vec{r}_{COP}) = \begin{Bmatrix} SL_{COP} \\ BL_{COP} \\ WL_{COP} \end{Bmatrix}
\end{equation}
The vector $\vec{r}_{CG,COP}$ is then simply the different between
both vectors.
\begin{equation}
\vec{r}_{CG,COP} = \vec{r}_{COP}-\vec{r}_{CG}
\end{equation}
The damping coefficient defined in equation (\ref{e:aeroM})
include $C_{l_p}$ which is the roll damping coefficient while $C_{m_q}$ is the pitch damping
coefficient. These coefficients are added which essentially inhibit
angular motion of the projectile. In addition, to these coefficients,
sometimes magnus coefficients are given as pure moments rather 
than forces acting at a distance. This can be given in the equation
below. 
\begin{equation}
M_{UA} = Qd (-C_{M\alpha}\frac{v}{V} + C_{N_{p\alpha}}\frac{w}{V}\frac{pd}{2V})
\end{equation}
Where $C_{M\alpha}$ replaces the moment produced by $C_{N\alpha}$ and
$C_{N_{p\alpha}}$ replaces the moment produced by
$C_{Y_{p\alpha}}$. It is possible to derive an equation between the
two different representations as given by the equations below.

\begin{equation}
\begin{matrix}
C_{M\alpha} = \frac{(SL_{COP}-SL_{CG})C_{N\alpha}}{d} \\
C_{N_{p\alpha}} = \frac{(SL_{MAG}-SL_{CG})C_{Y_{p\alpha}}}{d}
\end{matrix}
\end{equation}

\section{Stability and Control}

Controllability is formally stated as a system where any initial
state $x(0)=x_0$ and final state $x_1,t_1>0$, there exists a piecewise
continuous input $u(t)$ such that $x(t_1)=x_1$. 
For a fixed wing aircraft the system has 12 states with 8 dynamic
modes and 4 zero or rigid body modes. For a fixed wing aircraft the system has 12 states with 8 dynamic
modes and 4 zero or rigid body modes. A conventional aircraft has 4
controls to control these 12 modes. The easiest way to test the 
controllability of a system  is to compute the
controllability matrix. However, the controllability matrix must be
computed using a linearized model such that
$\dot{\vec{x}}=A\vec{x}+B\vec{u}$. In order to do this the aircraft
must be in equilibrium. For this example the aircraft is
set with an initial velocity of $20~m/s$ at an altitude of
$200~m$. The altitude command is set to $200~m$ and the heading
command is set to zero. Given the zero heading angle command and the
symmetry of the configurations investigated the rudder and aileron
commands are set to zero. Thus, only the thrust and elevator controls
are activated for the trimming procedure. Each configuration is
simulated for 200 seconds or until the derivatives of all states
except $\dot{x}$ are within a required tolerance. Using this
equilibrium point a linear model can be computed by using forward
finite differencing assuming that the
aircraft model is put in the form $\dot{\vec{x}} = F(\vec{x},\vec{u})$.
\begin{equation}
\dot{\vec{\delta x}} = \frac{F(\vec{x_0}+\Delta \vec{x_0},\vec{u_0})-F(\vec{x_0},\vec{u_0})}{\Delta
  \vec{x}}\vec{\delta x} + \frac{F(\vec{x_0},\vec{u_0}+\Delta
  \vec{u})-F(\vec{x_0},\vec{u_0})}{\Delta \vec{u}}\vec{\delta u}
\end{equation}
This linear model is the classic linear model where
$\dot{\vec{\delta x}}=A\vec{\delta{x}}+B\vec{\delta{u}}$. Using this linear model, the
controllability matrix can be computed as
\begin{equation}
W_C = [B~AB~A^2B~A^3B~...~A^{N-1}B]
\end{equation}
where N is the number of states in the system. With the controllability
matrix formulated, the rank of the matrix is computed. If the
$rank(W_C)=N$ the system is said to be controllable.

\section{PID Control}

For a conventional PID controller of an aircraft, the rudder, elevator
and aileron commands are set to 
\begin{equation}
\begin{matrix}
\delta_{r} =-K_v v \\
\delta_{e} = K_p(\theta-\theta_{c})+K_d{\dot \theta} \\ 
\delta_a = K_p(\phi-\phi_{c})+K_d{\dot \phi}
\end{matrix}
\end{equation}
The Euler angle commands $\phi_{c}$ and $\theta_{c}$ are set using the
following relationships:
\begin{equation}
  \begin{matrix}
  \phi_{c} = K_p(\psi-\psi_c)+K_d\dot{\psi} \\
  \theta_{c} = K_p(z-z_c) + K_d\dot{z} + K_{I}\int{z-z_c}dt
  \end{matrix}
\end{equation}
The control scheme defined above is a conventional inner loop-outer
loop control of a fixed wing aircraft using a PID tracking
controller.

\newpage

\section{AIRCRAFT NOMENCLATURE}
\begin{tabbing}
  XXXXXXXXXX \= \kill% this line sets tab stop
  $x_i,y_i,z_i$ \> components of the mass center position vector in the
  inertial frame of aircraft $i$ (m) \\
  $\phi_i,\theta_i,\psi_i$ \> Euler roll,pitch, and yaw of aircraft
  $i$ (rad) \\
  $u_i,v_i,w_i$ \> components of the mass center velocity vector in the
  body frame of aircraft $i$ (m/s) \\
  $p_i,q_i,r_i$ \> components of the mass center angular velocity vector in the
  body frame of aircraft $i$ (rad/s) \\
  ${\vec r}_{A\rightarrow B}$ \> position vector from a generic point A
  to a generic point B(m) \\
  ${\vec V}_{A/B}$ \> velocity vector of a generic point A with respect
  to frame B (m/s) \\
  $\textbf{T}_{AB}$ \> generic transformation matrix rotating a vector from
  the frame B to frame A \\
  $\textbf{H}_i$ \> relationship matrix of Euler angle derivatives to
  body angular velocity components of aircraft $i$ \\
  $m_i$ \> mass of aircraft $i$ (kg) \\
  $I_i$ \> moment of inertia matrix of aircraft $i$ taken about the mass center
  in the body frame($kg-m^2$) \\
  $X_i,Y_i,Z_i$ \> components of the total force applied to aircraft $i$ in
  body frame(N) \\
  $L_i,M_i,N_i$ \> components of the total moment applied to aircraft
  $i$ in body frame(N-m) \\
  $X_{Wi},Y_{Wi},Z_{Wi}$ \> total weight force applied to aircraft
  $i$ (N) \\
  $L,D$ \> Lift and Drag on Aircraft (N) - Not to be confused with Roll moment\\
  $g$ \> gravitational constant on Earth $(m/s^2)$ \\
  $\rho$ \> atmospheric density($kg/m^3$) \\
  $S_i$ \> reference area of wing on aircraft $i$ ($m^2$) \\
  $b_i$ \> Wingspan of aircraft $i$ (m) \\
  $\bar{c}_i$ \> mean chord of wing on aircraft $i$ (m) \\
  $\alpha$ \> Angle of attack (rad) \\
  $\beta$ \> Slideslip angle (rad) \\
  $C_L,C_D,C_m$ \> Lift, Drag and Pitch Moment coefficients \\
  $\delta_t,\delta_a,\delta_r,\delta_e$ \> thrust, aileron, rudder,
  and elevator control inputs(rad) \\
  $S_B(\vec{r})$ \> skew symmetric matrix operator on a vector
  expressed in the body frame. \\
  $K_p,K_d,K_I$ \> proportional, derivative, and integral control
  gains\\
  $V$ \> Total airspeed (m/s) \\
  $\hat{p},\hat{q},\hat{r}$ \> Non-dimensional angular velocities \\
  $l$ \> Distance from center of mass to aerodynamic center of the
  tail (m) \\
  $l_t$ \> Distance from aerodynamic center of main wing to
  aerodynamic center of tail (m) \\
  $\alpha_0$ \> zero lift angle of attack (rad) \\
  $C_{L0}$ \> Zero angle of attack lift coefficient \\
  $C_{m\alpha}$ \> Pitch moment curve slope versus $\alpha$ \\
  $C_{L\alpha}$ \> Lift curve slope \\
  $C_{mq}$ \> Pitch damping coefficient \\
  $C_{m\delta_e}$ \> Pitch moment curve slope versus elevator
  deflection angle \\
  $a_{\infty}$ \> Speed of sound (m/s) \\
  $\mu_{\infty}$ \> Viscosity of Fluid $kg/(m-s)$ \\
  
\end{tabbing}

\newpage

\section{EQUATIONS}

\begin{multicols}{2}

\noindent Mach Number and Reynolds Number

\beq
\beqn
M_{\infty} = \frac{V}{a_{\infty}} \\
\ \\
Re = \frac{\rho V \bar{c}}{\mu_{\infty}}
\eeqn
\eeq

\noindent Total Velocity

\beq
V = \sqrt{u^2 + v^2 + w^2}
\eeq

\noindent Angle of Attack and Sideslip

\beq
\beqn
\alpha = tan^{-1}\left(\frac{w}{u}\right)\\
\beta = sin^{-1}\left(\frac{v}{V}\right)
\eeqn
\eeq

\noindent Lift Drag and Moment

\beq
\beqn
Lift~(L) = \frac{1}{2} \rho V^2 S C_L \\
Drag~(D) = \frac{1}{2} \rho V^2 S C_D\\
Roll~Moment~(L) = \frac{1}{2} \rho V^2 S b C_l\\
Pitch~Moment~(M) = \frac{1}{2} \rho V^2 S \bar{c}C_m\\
Yaw~Moment~(N) = \frac{1}{2} \rho V^2 S b C_n\\
\eeqn
\eeq

\noindent Lift and Drag Coefficients

\beq
\beqn
C_L = C_{L0} + C_{L\alpha}\alpha\\
C_L = C_{L\alpha}(\alpha-\alpha_0)\\
C_D = C_{D0} + C_{D\alpha}\alpha^2 \\
C_D = C_{D0} + k{C_L}^2 \\
\eeqn
\eeq

\noindent Non-dimensional Angular velocities

\beq
\beqn
\hat{p} = pb/2V\\
\hat{q} = q\bar{c}/2V\\
\hat{r} = rb/2V
\eeqn
\eeq

\noindent Pitch Moment equation

\beq
C_m = C_{m0} + C_{m\alpha}\alpha + C_{m\delta_e}\delta_e +
C_{mq}\hat{q}
\eeq

\beq
\beqn
C_{m0} = C_{MAC} + C_{L0}\bar{x}_{sm} \\
\bar{x}_{sm} = \frac{x_{cg}}{\bar{c}} - \frac{x_{acW}}{\bar{c}} \\
C_{m\alpha} = \left(C_{L\alpha,W} +
\frac{S_t}{S}C_{L\alpha,t}\right)\bar{x}_{sm} - V_HC_{L\alpha,t}\\
V_H = \frac{l_tS_t}{S\bar{c}} \\
C_{m\delta_e} = \left(C_{Lt\delta_e}\frac{S_t}{S}\right)\bar{x}_{sm}-V_HC_{Lt\delta_e}\\
C_{mq} = 2C_{L\alpha t}\frac{l^2}{\bar{c}^2}
\eeqn
\eeq

\noindent Max Lift to Drag Ratio (Only valid if ${C_{L0}=0}$)

\beq
\alpha_{max,L/D} = \sqrt{\frac{C_{D0}}{C_{D\alpha}}}
\eeq

\noindent Lift to Drag when $T=0$ (Sum of Forces still zero)

\beq
\beqn
\frac{D}{L} = tan(\alpha)\\ %%%Where does this come from? Sum of Forces = 0
Lcos(\alpha) + Dsin(\alpha) = W
\eeqn
\eeq

\noindent Airfoil and Wing Aerodynamics

\beq
\beqn
x_{ac} = c/4 & 
a = \frac{a_0}{1+\frac{a_0}{\pi e AR}} \\
AR = \frac{b^2}{S}
\eeqn
\eeq


\noindent Standard Atmosphere

\beq
\beqn
\rho = 1.225~kg/m^3 = 0.00238~slugs/ft^3\\
\mu_{\infty} = 1.81x10^{-5}~kg/(m-s)\\
a_{\infty} = 331.3~m/s
\eeqn
\eeq

\noindent General Notes

\begin{enumerate}
  \item In trim or steady and level or cruise $q=0$, $C_m=0$,$L=W$, $T=D$
  \item For symmetric airfoil $C_{MAC} = 0$ and
    $C_{L0}=0$ thus $C_{m0} = 0$
  \item For a flat plate all symmetric properties apply but
    $a_0 = 2\pi$
  \item Tail surfaces are always assumed to be flat plates
  \item For longitudinal problems, $\beta = 0$ so $v=0$ (side velocity)
\end{enumerate}

\end{multicols}

\end{document}
