\subsection{Driving Lab Experiment}

Your task is to team with two other lab partners and go for a ride
around the town or get some coffee or just run an errand. Here are the rules.

\begin{enumerate}

\item The driver is responsible for one thing and that is driving. Do
  not take your eyes off the road or text while driving. You must be
  safe at all times and get all of your passengers to and from your
  destination safely. For this assignment the driver must own a
  vehicle with a functioning odometer and speedometer.  I would
  recommend resetting your trip meter. If your car does not have a
  trip meter you must write down the initial mileage of your vehicle
  before you depart. 

\item Passenger number 1. Your task is to read off the drivers speed
  every 10 seconds on the departure leg and every 30 seconds on the
  return leg. I would bring a pencil and paper or your computer with
  excel open and write down the time in intervals of 10 seconds on the
  departure leg and 30 seconds for the return leg. While the car is
  driving you can then write down the speed by looking at the
  speedometer and writing the speed in mph. I would recommend
  converting mph to miles/second so that your units will match up. 

\item Passenger number 2 is responsible for measuring the distance
  traveled in units of 0.1 miles using the odometer. I would recommend
  synchronizing watches with passenger number 1 and passenger number 2
  so that your time vectors line up. Your table should have time in
  one column and position in increments of 0.1 miles in another
  column.

\end{enumerate}

Once all of this data is compiled you need to generate the following plots.

\begin{enumerate}

\item Velocity versus time

\item Position versus time

\item Use first order differentiation and compute velocity versus time
  using part b. Plot this alongside part a and compare your
  results. For the down leg and the return leg. Which is more
  accurate? 

\item Use the trapezoidal rule to compute position versus time using
  part a and plot it alongside part b. Compare your results with the
  down leg and the return leg. Which is more accurate? 

NOTE ALL THREE OF YOU MUST COMPLETE THE ASSIGNMENT. YOU MAY
TAKE DATA AND HAVE FUN TOGETHER BUT EVERYONE MUST STILL COMPLETE THE
LAB.  

\end{enumerate}

