\section{Notes for Next Semester}

\begin{enumerate}

\item The final project next semester should have some sort of design
  component or basically something they have to solve that does not
  have an analytical solution.

\item Change wording in syllabus so students don't think they have to
  email me everytime they come to my office

\item Allow 2 resubmissions on Sakai - it's starting to happen alot.

\item Teach Taylor Series, Euler's Method and Search Methods before
  the first test. As such you will only have two tests during the
  semester.

\item Add Smartbook http://connect.mheducation.com/ to the syllabus as
  an option for the textbook. ISBN: 9781259238048

\item Ask all students for schedule of events so it doesn't overlap
  with important events

\item If you fail my final you fail the class

\end{enumerate}

\section{Schedule (With Class Number)}

\begin{enumerate}

\item Monday - 1/12 

  Class: Go through survey from last semester. Have
  students go over syllabus themselves and ask questions after 10
  minutes. Go over Sakai tool and computers, Talk about youtube
  videos. Notify students of first homework assignment due on
  Friday(screenshot of computer in a WORD Document converted to
  PDF). Notify students that they must have MATLAB by 5th
  class(Homework 2 - Screenshot again). Then stay in groups and
  complete survey (using polleverywhere.com).  
  
  Homework: Watch Uploading/Downloading Assignments if they are
  having trouble with Homework 1

\item Wednesday - 1/14 - Engineering Methods, Approximations and Round
  Off Errors

  Class: Lecture on Chapter 1

  Homework: Complete Hw 1

\item Friday - 1/16 - Approximations and Round Off Errors - Hw 1 Due

  Class: Lecture on Chapter 3

  Homework: Make sure to have MATLAB installed and watch all youtube
  videos from Palm Chapter 1

\item Monday - 1/19 - Holiday

\item Wednesday - 1/21 - Overview of MATLAB - Hw 2 Due

  Class: Work on Homework 3
  
  Homework: Watch videos on Arrays Palm Chapters 2 and 3

\item Friday - 1/23 - Arrays

  Class: Work on homework 3

  Homework: Watch videos on Functions 

\item Monday - 1/26 - Functions

  Class: Work on homework 3 in class

  Homework: Watch videos on Conditionals and If/else/end

\item Wednesday - 1/28 - Conditionals and If/else/end

  Homework 3 due

  Class: Work on Homework 4

  Homework: Watch videos on loops

\item Friday - 1/30 - Loops

  Class: Work on Homework 4

  Homework: Watch videos on 2D and 3D plotting

\item Monday - 2/2 - Plotting 

  Class: Work on Homework 4

  Homework: Watch videos on Meshing

\item Wednesday - 2/4 - Meshing

  Homework 4 Due

  Class: Work on Homework 5
  
  Homework: Watch videos on Animating

\item Friday - 2/6 - Animating

  Class: Work on Homework 5

  Homework: Work on Homework 5

\item 2/9 - Monday - Review
  
  Homework 5 Due

\item 2/11 - Wednesday - Test 1

-----------------------------------------------------------

\item Friday - 2/13 - Finite Differencing and Taylor Series (Computation of Error)

  Homework 6 Assigned 

\item Monday - 2/16 - Go over homework 4 and Test 1

\item Wednesday - 2/18 - Bracketing Methods with Examples

\item Friday - 2/20 - Open Methods and One Dimensional Optimization

\item Monday - 2/23 - MATLAB Day 

\item Wednesday - 2/25 - Review - Homework 6 Due

\item Friday - 2/27 - Test 2 - (Assign Project)

\item 3/2-3/6 - Spring Break 

\item Monday - 3/9 - (Go over project,Test 2, Hw 6) - 1

  Assign HW 7

\item Wednesday - 3/11 - Gaussian Elimination, Matrix Inverse, Determinants and
  Eigenvalue Analysis - 2

\item Friday - 3/13 - Example Problems/MATLAB Day - 3

\item 3/16 - Linear and Polynomial Regression - Monday - 4

\item 3/18 - Example Problems - HW 7 Due - Wednesday - 5
 
\item 3/20 - MATLAB Day - Friday - 6

\item 3/23 - Linear and Polynomial Interpolation, Linear and Quadratic Splines -
  Monday - 7

\item 3/25 - Example Problems - Wednesday - 8

\item 3/27 - MATLAB Day - Friday - 9

\item 3/30 - 2-D Interpolation, Fourier Series Analysis - Monday - 10

\item 4/1 - Example Problems - April Fools joke - Wednesday - Pop Quiz
  - 11

\item 4/3 - Review - Friday - 12

\item 4/6 - Test 3 - Monday

\item 4/8 - Newton-Cotes Integration - Assign Hw - Wed

\item 4/10 - Example Problems - Fri

\item 4/13 - MATLAB Day - Mon - Have Cody proctor class and help
  students with code

\item 4/15 - Wed - Assign Hw - Heun's Method and Runge-Kutta-4

\item 4/17 - Example Problems

\item 4/20 - MATLAB Day 

\item 4/22 - Work on Group Projects

\item 4/24 - Fri - Presentations?

\item 4/27 - Mon - Presentations? 

\item 4/29 - Wed - MATLAB Jeopardy

\item 5/1 - Fri - Review

\end{enumerate}

